---
title: "Budweiser"
author: "Bernard Strzalkowski and Keith Sikes"
date: '2022-06-18'
output: html_document
---

```{r map graphics}
if (!require("usmap", quietly = TRUE)) install.packages("usmap")
install.packages("BiocManager")

library(usmap)
library(stringr)
library(ggplot2)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if("e1071" %in% rownames(installed.packages()) == FALSE) {
  install.packages("e1071")}
if("caTools" %in% rownames(installed.packages()) == FALSE) {
  install.packages("caTools")}
if("caret" %in% rownames(installed.packages()) == FALSE) {
  install.packages("caret")}
  if("moments" %in% rownames(installed.packages()) == FALSE) {
  install.packages("moments")}
    
library(readr)
library(dplyr)
library(tidyr)
require(e1071)
require(caTools)
require(caret)
library(class)

```

## Basics 
### Input, group_by, sum, change column names, merge

1.	How many breweries are present in each state?
2.	Merge beer data with the breweries data. Print the first 6how bservations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)

```{r}
setwd("C:/Users/Berna/RCoding/Budweiser/Budweiser")
Breweries <- read_csv("Breweries.csv")
Beers <- read_csv("Beers.csv", col_types = cols(Beer_ID = col_integer(), 
    Brewery_id = col_integer()), na = "NA")

# rename columns
rename(Beers, Brew_ID = Brewery_id,Beer=Name) -> Beers
rename(Breweries, state = State,Brewery=Name) -> Breweries

# group by state, count breweries in state, order by count and state.
Q1 <- Breweries %>% group_by(state) %>% summarise(Density = n()) %>% arrange(desc(Density),state)

# merge many to one 
Q2 <- inner_join(Beers,Breweries,by ="Brew_ID")

head(Q2)
tail(Q2)

write_csv(Q1,file="StateBrew.csv")
```

one can use Excel to format the data into a 6 column table for easy display

## Choropleth

> Data is dataframe and must have state code and numerical value
size = line size, values is column with data to use for colors
name = name of label and label defines type of label. 
labs = label or title.

```{r}
plot_usmap(regions ="state", data = Q1, size=.2, values = "Density") + 
    scale_fill_continuous( low = "pink", high = "darkred", 
                         name = "Breweries", label = scales::comma) +
  theme(legend.position = "right") +
  labs(title = "Brewery Density by State") 
#+   theme(panel.background = element_rect(color = "grey90", fill = "lightblue")) 
```




### Median ABV
```{r}
medianABV = median(Q2$ABV)
medianABV

medianABVna = median(Q2$ABV, na.rm = TRUE)
medianABVna

Q2 %>% group_by(state) %>% 
  summarize(Median_ABV = median(ABV,na.rm=TRUE),
            Max_ABV = max(ABV, na.rm=TRUE)) -> StateABV
StateABV

# sorting
Q2 %>% arrange(desc(ABV))-> Temp 
head(Temp,1)

# barcharts (use reorder to change the order of the state (as factor) to match maxstateABV)
d <- ggplot(StateABV, aes(y=Median_ABV,x=reorder(as.factor(state),Median_ABV), fill = Median_ABV))
d <- d + geom_col(stat="identity") + coord_flip() +
  ggtitle("Median Alcohol By Volume Content By State") +
  xlab("State") +
  ylab("ABV") +
  theme(axis.text = element_text(colour = "black", size = 6)) +
  theme(panel.background = element_rect(fill = "white"))
d

```

### Median IBU

```{r}
medianIBU = median(Q2$IBU)
medianIBU

medianIBUna = median(Q2$IBU, na.rm = TRUE)
medianIBUna

Q2 %>% group_by(state) %>% 
  summarize(Median_IBU = median(IBU,na.rm=TRUE),
            Max_IBU = max(IBU, na.rm=TRUE)) -> StateIBU
StateIBU

# sorting
Q2 %>% arrange(desc(IBU))-> Temp 
head(Temp,1)

# barcharts (use reorder to change the order of the state (as factor) to match Median_ABV)
d <- ggplot(StateIBU, aes(y=Median_IBU,x=reorder(as.factor(state), Median_IBU), fill = Median_IBU))
d <- d + geom_col(stat="identity") + coord_flip() +
  ggtitle("Median International Bitterness Unit By State") +
  xlab("State") +
  ylab("IBU") +
  theme(axis.text = element_text(colour = "black", size = 6)) +
  theme(panel.background = element_rect(fill = "white"))
d

```

### Max ABV
```{r}
medianABV = median(Q2$ABV)
medianABV

medianABVna = median(Q2$ABV, na.rm = TRUE)
medianABVna

Q2 %>% group_by(state) %>% 
  summarize(Median_ABV = median(ABV,na.rm=TRUE),
            Max_ABV = max(ABV, na.rm=TRUE)) -> StateABV
StateABV

# sorting
Q2 %>% arrange(desc(IBU))-> Temp 
head(Temp,1)

# barcharts (use reorder to change the order of the state (as factor) to match maxstateABV)
d <- ggplot(StateABV, aes(y=Max_ABV,x=reorder(as.factor(state),Max_ABV), fill = Max_ABV))
d <- d + geom_col(stat="identity") + coord_flip() +
  ggtitle("Max Alcohol By Volume Content By State") +
  xlab("State") +
  ylab("ABV") +
  theme(axis.text = element_text(colour = "black", size = 6)) +
  theme(panel.background = element_rect(fill = "white"))
d

```
```{r}
medianIBU = median(Q2$IBU)
medianIBU

medianIBUna = median(Q2$IBU, na.rm = TRUE)
medianIBUna

Q2 %>% group_by(state) %>% 
  summarize(Median_IBU = median(IBU,na.rm=TRUE),
            Max_IBU = max(IBU, na.rm=TRUE)) -> StateIBU
StateABV

# sorting
Q2 %>% arrange(desc(IBU))-> Temp 
head(Temp,1)

# barcharts (use reorder to change the order of the state (as factor) to match maxstateABV)
d <- ggplot(StateIBU, aes(y=Max_IBU,x=reorder(as.factor(state),Max_IBU), fill = Max_IBU))
d <- d + geom_col(stat="identity") + coord_flip() +
  ggtitle("Max International Bitterness Unit By State") +
  xlab("State") +
  ylab("IBU") +
  theme(axis.text = element_text(colour = "black", size = 6)) +
  theme(panel.background = element_rect(fill = "white"))
d

```

#### Comparing ABV Data

```{r}
library(quantmod)
library(ggplot2)
library(TTR)
library(ggrepel)

#plot(x = StateABV$Max_ABV, y = StateABV$Median_ABV, col = as.factor(StateABV$state))

# df <- ggplot(data = StateABV, aes(x = Max_ABV, y = Median_ABV, fill = as.factor(state)) + geom_point()

# Calculate Moving Average for Median
StateABV$MA5 <- TTR::SMA(StateABV$Median_ABV, n=5)

ggplot(data = StateABV, aes(x = Max_ABV, y = Median_ABV)) + 
  scale_color_gradient(low = "cyan", high = "blue4") +
  geom_point(aes(colour=Median_ABV)) +
  geom_smooth(method = "loess", colour = "darkblue") +
  ggtitle("Comparing Median to Max ABV") +
  xlab("Max ABV") +
  ylab("Median ABV") +
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.grid = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75))


```


```{r}
library(quantmod)
library(ggplot2)
library(TTR)
library(ggrepel)

#plot(x = StateABV$Max_ABV, y = StateABV$Median_ABV, col = as.factor(StateABV$state))

# df <- ggplot(data = StateABV, aes(x = Max_ABV, y = Median_ABV, fill = as.factor(state)) + geom_point()

# Calculate Moving Average for Median
StateABV$MA5 <- TTR::SMA(StateABV$Median_ABV, n=5)

ggplot(data = StateABV, aes(x = Median_ABV, y = Max_ABV)) + 
  scale_color_gradient(low = "cyan", high = "blue4") +
  geom_point(aes(colour=Max_ABV)) +
  geom_smooth(method = "loess", colour = "darkblue") +
  ggtitle("Comparing Max to Median ABV") +
  xlab("Median ABV") +
  ylab("Max ABV") +
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.grid = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75))


```

```{r}
library(quantmod)
library(ggplot2)
library(TTR)
library(ggrepel)
library(tidyverse)
library(tidyr)

#plot(x = StateABV$Max_ABV, y = StateABV$Median_ABV, col = as.factor(StateABV$state))

# df <- ggplot(data = StateABV, aes(x = Max_ABV, y = Median_ABV, fill = as.factor(state)) + geom_point()

BeerComparison <- Q2
BeerComparison <- BeerComparison[!is.na(BeerComparison$IBU),]

ggplot(data = BeerComparison, aes(x = ABV, y = IBU)) + 
  scale_color_gradient(low = "cyan", high = "blue4") +
  geom_point(aes(colour=IBU)) +
  geom_smooth(method = "loess", colour = "darkblue") +
  ggtitle("IBU to ABV Beer Comparison") +
  xlab("ABV") +
  ylab("IBU") +
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.grid = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75))


```

```{r}
library(quantmod)
library(ggplot2)
library(TTR)
library(ggrepel)
library(tidyverse)
library(tidyr)

#plot(x = StateABV$Max_ABV, y = StateABV$Median_ABV, col = as.factor(StateABV$state))

# df <- ggplot(data = StateABV, aes(x = Max_ABV, y = Median_ABV, fill = as.factor(state)) + geom_point()

BeerComparison <- Q2
BeerComparison <- BeerComparison[!is.na(BeerComparison$IBU),]

ggplot(data = BeerComparison, aes(x = IBU, y = ABV)) + 
  scale_color_gradient(low = "cyan", high = "blue4") +
  geom_point(aes(colour=ABV)) +
  geom_smooth(method = "loess", colour = "darkblue") +
  ggtitle("ABV to IBU Beer Comparison") +
  xlab("IBU") +
  ylab("ABV") +
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.grid = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75))


```

### filtering using text strings. 

>**grepl()**: grepl() function will is used to return the value TRUE if the specified string pattern is found in the vector and FALSE if it is not found.
Syntax: grepl(pattern, string, ignore.case=FALSE)

```{r}
list = grepl("ale",Q2$Style, ignore.case=TRUE)
alelist <- Q2 %>% filter(grepl("ale",Style, ignore.case=TRUE) & 
                           !(grepl("IPA",Style)|grepl("IPA",Beer)))
IPAlist <- Q2 %>% filter((grepl("IPA",Style)|grepl("IPA",Beer)))
alelist$type <- "ale"
IPAlist$type <- "IPA"
Beerlist <- rbind(alelist,IPAlist)
KNND <- as.factor(Beerlist$type)
KNNI <- Beerlist[,c(3,4)]
```

### Create datasets for KNN analysis (one data.frame with outcomes, a
second data.frame with the independent variables (ABV and IBU))



Beers[Beers$ABV>0.12&!is.na(Beers$ABV),]
Q2[grepl("Bud",Q2$Beer,ignore.case=TRUE),]

